---
title: "project1.2"
author: "Adit"
date: "August 24, 2014"
output: html_document
---
After loading the data, one can see that there are three variables with 60 observations.3 levels of dosage - 0.5,1,2 and two levels of suppliments - OJ(orange juice) and VC (Vitamin C).

```{r,echo=FALSE,warning=FALSE,message=F}
library(ggplot2)
library(gridExtra)
library(plyr)
library(lattice)
library(reshape2)

#loading the data file & undertanding dataset
data<-ToothGrowth
summary<-str(data) #three variables with 60 observations.
#factor(data$dose) # 3 levels of dosage - 0.5,1,2
#factor(data$supp) # two levels of suppliments - OJ(orange juice) and VC (Vitamin C)

Dose<-factor(data$dose)
```




Below we create a boxplot and a violin plot to understand the variablity of the length based on the suppliment and dosage.

```{r,echo=FALSE,warning=FALSE,message=F}
b1<-ggplot(data=data,aes(y=len,x=supp,fill=Dose))+
  geom_boxplot()+facet_grid(.~dose) +
  ggtitle("Boxplot of Tooth Length w.r.t Suppliments")+
  xlab("Suppliments")+
  ylab("Tooth Length") +
  theme(axis.title=element_text(size=8),plot.title=element_text(size=8,face="bold"))

v1<-ggplot(data=data,aes(y=len,x=factor(dose),fill=supp))+
  geom_violin()+facet_grid(.~supp) +  
  ggtitle("Violin of Tooth Length w.r.t Suppliments and Dosage")+  
  xlab("Dosage")+  ylab("Tooth Length")+
  theme(axis.title=element_text(size=8),plot.title=element_text(size=8,face="bold"))
```

```{r,echo=TRUE,warning=FALSE,message=F}
grid.arrange(b1,v1,ncol=2)
```

One can see that there is high variablity in the Tooth length based on suppliments and dosage. The potential variations are 15 (6C2). 

We use the t test to reject the 15 null hypothesis that:
<br>1. As the dosage increases the Tooth Length decereases
<br>2. For the same dosage, VC provides a bigger increase in Tooth Length as compared to OJ.

```{r, echo=FALSE,warning=F,message=FALSE}
#breaking the data into different dosages
test1<-subset(ToothGrowth,ToothGrowth$dose==0.5)
test2<-subset(ToothGrowth,ToothGrowth$dose==1)
test3<-subset(ToothGrowth,ToothGrowth$dose==2)

#Performing multiple t tests for the following variations. 
#In effect we have 15 variations 6C2
```

I have shown a few of the 15 tests below:

```{r, echo=TRUE,warning=F,message=FALSE}
#1. OJ@0.5 w.r.t OJ@1
h1<-t.test(test1[test1$supp=='OJ',]$len,test2[test2$supp=='OJ',]$len,paired = FALSE,var.equal = TRUE)
#2. OJ@0.5 w.r.t OJ@2
h2<-t.test(test1[test1$supp=='OJ',]$len,test3[test3$supp=='OJ',]$len,paired = FALSE,var.equal = TRUE)
#3. OJ@1 w.r.t OJ@2 almost the same
h3<-t.test(test2[test2$supp=='OJ',]$len,test3[test3$supp=='OJ',]$len,paired = FALSE,var.equal = TRUE)

#4. VC@0.5 w.r.t VC@1
h4<-t.test(test1[test1$supp=='VC',]$len,test2[test2$supp=='VC',]$len,paired = FALSE,var.equal = TRUE)
#5. VC@0.5 w.r.t VC@2
h5<-t.test(test1[test1$supp=='VC',]$len,test3[test3$supp=='VC',]$len,paired = FALSE,var.equal = TRUE)
#6. VC@1 w.r.t VC@2
h6<-t.test(test2[test2$supp=='VC',]$len,test3[test3$supp=='VC',]$len,paired = FALSE,var.equal = TRUE)

#7. OJ@0.5 w.r.t VC@0.5
h7<-t.test(test1[test1$supp=='OJ',]$len,test1[test1$supp=='VC',]$len,paired = FALSE,var.equal = TRUE)
#8. OJ@0.5 wr.t VC@1 almost the same
h8<-t.test(test1[test1$supp=='OJ',]$len,test2[test2$supp=='VC',]$len,paired = FALSE,var.equal = TRUE)
#9. OJ@0.5 wr.t VC@2
h9<-t.test(test1[test1$supp=='OJ',]$len,test3[test3$supp=='VC',]$len,paired = FALSE,var.equal = TRUE)
```

```{r, echo=FALSE,warning=F,message=FALSE}
#10. OJ@1 w.r.t VC@0.5
h10<-t.test(test2[test2$supp=='OJ',]$len,test1[test1$supp=='VC',]$len,paired = FALSE,var.equal = TRUE)
#11. OJ@1 w.r.t VC@1
h11<-t.test(test2[test2$supp=='OJ',]$len,test2[test2$supp=='VC',]$len,paired = FALSE,var.equal = TRUE)
#12. OJ@1 w.r.t VC@2 the same
h12<-t.test(test2[test2$supp=='OJ',]$len,test3[test3$supp=='VC',]$len,paired = FALSE,var.equal = TRUE)

#13. OJ@2 w.r.t VC@0.5
h13<-t.test(test2[test2$supp=='OJ',]$len,test1[test1$supp=='VC',]$len,paired = FALSE,var.equal = TRUE)
#14. OJ@2 w.r.t VC@1
h14<-t.test(test2[test2$supp=='OJ',]$len,test2[test2$supp=='VC',]$len,paired = FALSE,var.equal = TRUE)
#15. OJ@2 w.r.t VC@2 #the same
h15<-t.test(test2[test2$supp=='OJ',]$len,test3[test3$supp=='VC',]$len,paired = FALSE,var.equal = TRUE)
```

