---
title: "Project1.1"
author: "Adit"
date: "August 17, 2014"
output: html_document
---

```{r,echo=TRUE,warning=FALSE,message=FALSE}
#library
library(ggplot2)
library(gridExtra)

#initializing the variables
lambda<-0.2
samples<-40
trials<-1000


sampleDist<-0 #exponential distribution of every 40 samples
meanSample<-0 #mean of the 40 samples
dist<-0 #saving all the distributions over 1000 trials
mean<-0 #saving all the means from teh 1000 trials.

for (i in 1:trials){
  sampleDist<-rexp(samples,lambda)
  meanSample<-mean(sampleDist)
  dist<-c(dist,sampleDist)
  mean<-c(mean,meanSample)
}
mean<-data.frame(mean=mean[-1])
dist<-data.frame(dist=dist[-1])

#comparing means 
meanT<-1/lambda
meanD<-colMeans(mean)

#comparing the sds
sdT<-1/(lambda*sqrt(samples))
sdD<-sd(mean$mean)

#finding the 95% interval about the mean
range<-round(c((qnorm(.975)*sdT*c(-1,1)+meanT),as.numeric(meanT)),2)

#checking to see % of means of the 1000 trials within 95% of the population mean.
confidence<-100*(sum(mean$mean<range[2] & mean$mean>range[1]))/trials
```


```{r,echo=FALSE,warning=FALSE}
#plot of the exponential distribution
p<-ggplot(data=dist,aes(dist))+
  geom_histogram(aes(y=..density..),binwidth=.3)+
  scale_x_continuous(limits=c(0,40))+
  xlab("sample values(10K samples)") +
  ylab("probability") + 
  ggtitle("Density histogram of exponential distribution")+
  theme(axis.title=element_text(size=8),plot.title=element_text(size=8,face="bold"))


#histogram of means
h<-ggplot(data=mean,aes(x=mean)) +
  geom_histogram(aes(y=..density..),binwidth=.2,colour="black", fill="white") +
  geom_density(alpha=.4, fill="red",size=1) +
  xlab("mean values of each trial") +
  ylab("probability") + 
  ggtitle("Density histogram of means of each trial")+
  geom_vline(xintercept=range, linetype="dotted",color='blue')+
  theme(axis.title=element_text(size=8),plot.title=element_text(size=8,face="bold"))+
  scale_x_continuous(breaks=range,labels=c(range[1],range[2],range[3]))
```
<br>
<br>
<br>
The following are plots explaing the exponential distribution (plot1) and the distribution of means of the sampling mean (plot2)

40 samples of the exponential distribution are pooled 1000 times. Plot 2 shows the mean of each of these pooled trials with the blue dotted lines indicating the 95% interval.

I have not show the code for the graphs since it would not fit in one page.

```{r, fig.align='center',fig.env='marginfigure',echo=FALSE,warning=FALSE}
grid.arrange(p,h,ncol=2)
```

<ol>
<li>Show where the distribution is centered at and compares it to the theoretical center of the distribution.
<br>As shown above in plot 2, the distribution is centered around `r meanD` while the theoratical center is at `r meanT`. </li>

<br><li>Show how variable it is and compare it to the theoretical variance of the distribution.
<br>As shown above in plot 2, the distribution has a standard deviation of `r sdD`, while the theoratical standard deviation is `r sdT`. </li>

<br><li> Show that the distribution is approximately normal. 
<br>As shown in plot 2, the distirbution of the means is approximately normal with 95% of the readings within 2 standard deviations of the mean. There are `r confidence`% of means of the 1000 trials within 2 standard deviations of the population mean </li>

<br><li> Evaluate the coverage of the confidence interval for 1/lambda: X¯±1.96Sn√n. 
<br>From plot 2 one can see the 95% confidence interval points as `r range[1]` and `r range[2]`. This suggests that the population mean (mean of the exponential distribution) with a 95% probabilty if within the range defined above.
